title: "Brownlows Bets"
author: Tristyn Thompson
date: 13 November 2017
output: html_document: default
 ---

<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 1137 times so far. Where do you think his profits are? Discuss this in a markdown document

# Methods

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 1137 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->
We are going to run a simulation using two functions, buildDY and the table function. The buildDY is a function that identifies the class size and a list that covers the length of the data; in this case it is a class of 35 and a list of the months in the year. The table function helps us to identify how many students share a birth month. The max duplicate (maxdup) helps us identify the max amount of people that shared the same birth month.
```{r}
buildDY <- function(csize=35){
  bdlist=1:365
  out <- sample(bdlist,csize,replace=T)
}

```

```{r}
try1=buildDY(csize = 35)
table(try1)

```

```{r}
maxdup = function(inlist){
    out <- max(table(inlist))
}

```




#Profit Estimate

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting!-->

We want to run the simulation many times to get as accurate as possible. 30, 50, 60, and even 1000 are good numbers. The goal is to get the middle 95%. This simulation is going to be running 30 times to get as close to accurate as possible. 
```{r echo=FALSE} 
buildDY <- function(csize=35){
  bdlist=1:365
  out <- sample(bdlist,csize,replace=T)
}

```

```{r echo=FALSE}
try1=buildDY(csize = 35)
table(try1)

```

```{r echo=FALSE}
maxdup = function(inlist){
    out <- max(table(inlist))
}

```





# Confidence interval of Brownlows earning

<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows-->
```{r echo=FALSE}
buildDY <- function(csize=35){
  bdlist=1:365
  out <- sample(bdlist,csize,replace=T)
}

```



```{r echo=FALSE}
maxdup = function(inlist){
    out <- max(table(inlist))
}
```

```{r echo=FALSE}
bucket=numeric()
for (j in 1:30){
  


brownlow=0
trials = numeric()

for (i in 1:1137) {
  a<-buildDY(csize = 35)
  trials[i] = maxdup(a)
 
}

out <- ifelse(trials>=2,1,-2) 
bucket[j] <- sum(out)
}
```


```{r}
mean(bucket)
sd(bucket)
```

The confidence interval would be between 400 and 600, using the 68/95 rule. The standard deviation is close to 40. By following the standard deviation rule, two down and two up, the range is from 35 to 39.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

Based on our simulations, we learn that Professor Brownlow is consistently making money. To have his balance at zero profit we will change the class size but keep it at a 2:1 bet. Through the simulations, we learned that the class size needs to be around 28-29 students. In this particular simulation, the class size is 29 and Professor Brownlow's profit is lower than when he had a class size of 35. 
```{r echo=FALSE}
buildDY <- function(csize=35){
  bdlist=1:365
  out <- sample(bdlist,csize,replace=T)
}

```


```{r echo=FALSE}
maxdup = function(inlist){
    out <- max(table(inlist))
}
```


```{r}
bucket=numeric()
for (j in 1:10){
  


brownlow=0
trials = numeric()

for (i in 1:1137) {
  a<-buildDY(csize = 29)
  trials[i] = maxdup(a)
 
}

out <- ifelse(trials>=2,1,-2) 
bucket[j] <- sum(out)
}
```
